---
title: "Part one"
author: "Maggie Walters"
date: "August 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data <- read.csv("spm.data.long.csv")
data <- subset(data, data$CONCLUSION == "CLINICAL" | data$CONCLUSION == "CONFIRMED")
county_vec <- unique(data$COUNTY)
install.packages("ggplot2")
```

## Determine the distribution of outbreak sizes
```{r Outbreak sizes}
outbreakSizeVector <- rep(NA, length(county_vec))
names(outbreakSizeVector) <- county_vec
for(i in 1:length(county_vec)){
  outbreakSizeVector[i] <- length(which(data$COUNTY == county_vec[i]))
}

#Quantile vales of outbreak sizes
quantile(outbreakSizeVector)
which.smallOutbreak <- which(outbreakSizeVector < 2)

#make status mat
status_mat <- matrix(rep(NA, length(county_vec) * 53), ncol = length(county_vec))
colnames(status_mat) <- county_vec
rownames(status_mat) <- seq(1:53)
for(i in 1:length(county_vec)){
  for(j in 1:53){
    x <- subset(data, data$COUNTY == county_vec[i] & data$WEEK == j)
    status_mat[j,i] <- nrow(x)
  }
}
status_mat <- as.data.frame(status_mat)

#parse down status mat to only those with more than one outbreak
status_mat <- status_mat[,-which.smallOutbreak]

#Find those with two weeks infection free after an outbreak
binary_status_mat <- matrix(ncol = dim(status_mat)[2], nrow = dim(status_mat)[1], NA)
for(i in 1:dim(status_mat)[2]){
  for(j in 1:dim(status_mat)[1])
  if(status_mat[j,i] != 0){
    binary_status_mat[j,i] <- 1
  }else{
    binary_status_mat[j,i] <- 0
  }
}

reSusceptible_counties <- binary_status_mat
for(i in 1:dim(status_mat)[2]){
  for(j in 1:50){
    if(binary_status_mat[j,i] == 1 & binary_status_mat[j+1,i] == 0 & binary_status_mat[j+2,i] == 0 & binary_status_mat[j+3,i] == 0){
      reSusceptible_counties[j+3,i] <- "RESUSCEPTIBLE"
    }
  }
}

#fill in susceptible for inital zeros
for(i in 1:dim(status_mat)[2]){
  first_case <- which(reSusceptible_counties[,i] == 1)[1]
  reSusceptible_counties[1:first_case - 1, i] <- "SUSCEPTIBLE"
}

for(i in 1:dim(status_mat)[2]){
  for(j in 3:50){
if(reSusceptible_counties[j,i] == 0 & reSusceptible_counties[j-1, i] != 1 & reSusceptible_counties[j-2, i] != 1){
  reSusceptible_counties[j,i] <- "RESUSCEPTIBLE"
}
    print(i)
    print(j)
}}

```


## Summary of Outbreak sizes chunk

All counties with only one case can definitely be considered susceptible after their initial infection. `r length(which.smallOutbreak)` counties were found to only have one case. 

For counties with more than one case, if a county has not experienced infeciton in the previous two weeks, that county is considered "RESUSCEPTIBLE". 

To visualize this, a graph comparing infectious periods and susceptible periods will be created.

```{r Graphics pre-work}
#infectious period
infectious_period <- matrix(nrow = length(names(status_mat)), ncol = 53)
row.names(infectious_period) <- names(status_mat)
for(i in 1:283){
  infectious_period[i,] <- c(which(reSusceptible_counties[,i] == 1 | 
                                     reSusceptible_counties[,i] == 0), 
                             rep(NA, 53 - length(which(reSusceptible_counties[,i] == 1 | 
                                                         reSusceptible_counties[,i] == 0))))
}
as.data.frame(infectious_period)

infectious_period <- t(infectious_period)
infectious_period <- as.vector(infectious_period)
infectious_period_new <- matrix(infectious_period, nrow = length(infectious_period), ncol = 2)
infectious_period_new[1:53,2] <- rep(names(status_mat)[1], 53)
for(i in 2:283){
  x <- (i - 1) * 53
  infectious_period_new[x + (1 : 53),2] <- rep(names(status_mat)[i], 53)
}
infectious_period_new <- as.data.frame(infectious_period_new)
remove <- which(is.na(infectious_period_new[,1]))
infectious_period_new <- infectious_period_new[-remove,]
infectious_period_new[,1] <- as.integer(infectious_period_new[,1])

#susceptible period
susceptible_period <- matrix(nrow = length(names(status_mat)), ncol = 53)
row.names(susceptible_period) <- names(status_mat)
for(i in 2:283){
  x <- which(reSusceptible_counties[,i] == "SUSCEPTIBLE" | reSusceptible_counties[,i] == "RESUSCEPTIBLE")
  susceptible_period[i,] <- c(x, rep(NA, 53 - length(x)))
}
as.data.frame(susceptible_period)

susceptible_period <- t(susceptible_period)
susceptible_period <- as.vector(susceptible_period)
susceptible_period_new <- matrix(susceptible_period, nrow = length(susceptible_period), ncol = 2)
susceptible_period_new[1:53,2] <- rep(names(status_mat)[1], 53)
for(i in 2:283){
  x <- (i - 1) * 53
  susceptible_period_new[x + (1 : 53),2] <- rep(names(status_mat)[i], 53)
}
susceptible_period_new <- as.data.frame(susceptible_period_new)
remove <- which(is.na(susceptible_period_new[,1]))
susceptible_period_new <- susceptible_period_new[-remove,]
susceptible_period_new[,1] <- as.integer(susceptible_period_new[,1])


```

```{r Graphics}
#horizontal line graph
qplot(infectious_period_new[,1], infectious_period_new[,2],
      data = infectious_period_new, 
      ylab = "COUNTY",
      xlab = "WEEK",
      geom = "line") 
#density plot
infectious_density_plot <- qplot(infectious_period_new[,1], 
      data = infectious_period_new, 
      ylab = "DENSTIY",
      xlab = "WEEK",
      main = "Infectious County Density by Week",
      geom = "density") 


#horizontal line graph
qplot(susceptible_period_new[,1], susceptible_period_new[,2],
      data = susceptible_period_new, 
      ylab = "COUNTY",
      xlab = "WEEK",
      geom = "line") 
#density plot
susceptible_density_plot <- qplot(susceptible_period_new[,1], 
      data = susceptible_period_new,
      ylab = "DENSTIY",
      xlab = "WEEK",
      main = "Susceptible County Density by Week",
      geom = "density")

```

```{r stats}
#infectious period
infectious_period <- matrix(nrow = length(names(status_mat)), ncol = 53)
row.names(infectious_period) <- names(status_mat)
for(i in 1:283){
  infectious_period[i,] <- c(which(reSusceptible_counties[,i] == 1 | 
                                     reSusceptible_counties[,i] == 0), 
                             rep(NA, 53 - length(which(reSusceptible_counties[,i] == 1 | 
                                                         reSusceptible_counties[,i] == 0))))
}
as.data.frame(infectious_period)

infectious_period_length <- matrix(NA, nrow = 283, ncol = 2)
infectious_period_length[,2] <- unique(names(status_mat))
for(i in 1:283){
  infectious_period_length[i,1] <- length(which(!is.na(infectious_period[i,])))
}
infectious_period_length <- as.data.frame(infectious_period_length)
infectious_period_length[,1] <- as.numeric(infectious_period_length[,1])

```

